### 2023-05-29 application.tex
up

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\iffalse
\alain{ me semble maintenant inutile
%
\begin{definition}{ (index)}. 
If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$  is upward-closed and $k \geq 0$, let $U_k= \bigcup_{0 \leq n \leq k} \pred^n(U)$.
The {\em index} $k(U)$ is the
smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$.
\end{definition}
%
%
If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set, the sets $\pred(U )$, and $\pred^{\geq k}(U )$ for
every $k \geq 0$ are upward-closed, thus Lemma~\ref{upward-closed stablizes} ensures the existence of $k(U)$.
%
\begin{remark}
$U^{k+1} $ can be rewritten $U^{k+1}= \bigcup_{0 \leq j \leq k+1} \pred^j(U) = 
U \cup \bigcup_{1 \leq j \leq k+1} \pred^j(U) =
U \cup \pred(\bigcup_{0 \leq j \leq k} \pred^j(U))
=  U \cup \pred(U^k)$.
\end{remark}

This ensures the following.

\begin{fact}\label{stop condition}
% Fact 4 (stop condition). 
If $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is an upward-closed set and $k \geq 0$ s.t. $U^k = U^{k+1}$ , then $U^\ell = U^k$ for all $\ell \geq k$, i.e.,
$k(U) \leq k$. This also implies that $\pred^*(U) = U^k$.
\end{fact}

\begin{lemma}
% Lemma 3 ([1]) 
% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,
 Given a basis of an upward-closed set $U \subseteq S$, and a state $s$ of an effective strongly WSTS, we can decide whether $U  \xrightarrow {*}{} s$.
%	we can reach $U$ from $s$.
\end{lemma}

\begin{proof}
We have to show that we can compute a basis of $U^{k+1}$ if we are given a basis of $U^k $. 
Then the
decidability of the stop condition follows directly. Let $B$ be a basis of $U^k$. 
We have
$$U^{k+1} = U \cup \pred(U^k ) = U \cup
\bigcup_{s' \in B}
\pred(\uparrow \{s' \}).$$

Since a finite basis of $\pred(\uparrow \{s' \}$) is computable for any $s'\in S$ by definition, we obtain a finite generating set of $U^{k+1}$ . By
Fact~\ref{fact basis}, we can compute a basis of $U^{k+1}$.
\end{proof}}

\fi
%%%%%%%%%%%%%%%%%%%

\section{Deciding resilience for extended automata}

\subsection{Pushdown automata}

%We may consider context-free grammars.

\begin{theorem}
The three {\sc resilience} problems are decidable for pushdown automata with $\Safe$ and $\Bad$ regular languages.
\end{theorem}

\begin{proof}
Recall that resilience is equivalent to  $\Bad \subseteq \pred^*(\Safe)$. For every pushdown automaton $A=(Q,\Sigma,...)$ and every regular language $L \subseteq Q \times \Sigma^*$, the set $\pred^*(L)$ is a computable regular language \cite{DBLP:journals/ipl/BouajjaniEFMRWW00}. Since $\Bad$ and $\Safe$ are both regular, we deduce that
%
%			Moreover, $post^*(\Bad)$ is also a computable regular language \alain{pourquoi ?}   $post^*(L)$ and
 the inclusion $\Bad \subseteq \pred^*(\Safe)$ is decidable and resilience too.
%
%The kRP is decidable: % let us compute $ \pred^{\leq k}(\Safe)$ (in polynomial time). 
Similarly, since $ \pred^{\leq k}(\Safe)$ is a computable regular language,
we deduce 
that the inclusion $\Bad \subseteq \pred^{\leq k}(\Safe)$ is decidable and $k$-resilience too.
Since $k$-resilience is decidable, bounded-resilience also is;
we can iteratively check whether $\Bad \subseteq \pred^{\leq k}(\Safe)$,
if this is the case, $k$-resilience holds, and we stop with a positive,
else we check whether $ \pred^{\leq k}(\Safe) =  \pred^{*}(\Safe)$,
if this is the case, then bounded-resilience do not hold,
else we continue.
\end{proof}
%
Let us remark that $\uparrow \post^*(q,w)$ is also a computable regular language, but we cannot directly apply previous results on WSTS since pushdown automata are not WSTS for the usual orderings (subword, prefix).
Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} for the subword ordering and $\uparrow \post^*$ is computable but context-free grammars are not strongly compatible.

%  \textcolor{red}{à faire Mathieu}

%
%

\subsection{Regular fifo automata}

Fifo automata and systems of communicating finite-state machines (CFSMs) are essentially finite automata that communicate through fifo channels (or queues). For the sake of simplicity, we consider automata with an unique fifo channel (that have the power of turing machines). In 1982, Pachl introduced in \cite{DBLP:journals/corr/cs-LO-0306121} the property for such systems to have a regular reachability set (or even a regular relation) and he showed that this gives a semi-algorithm to test non-reachability (with a regular inductive invariant to obtain a witness of non-reachability). Let $(q,w)$ be an initial state and $(q',w')$ be a state. When $\post^*(q,w)$ is known to be regular (even if we don't know how to compute it), we may enumerate all regular languages $L \subseteq Q \times \Sigma^*$ that contain $(q,w)$ and such that $\post(L) \subseteq L$ (inductive invariant). We know that if $(q',w') \not\in \post^*(q,w)$ then there exists, at least, a regular language $L$ such that $\post^*(q,w) \subseteq L$ and $(q',w') \not\in L$, since $\post^*(q,w)$ is such inductive invariant. This strategy provides a semi-algorithm for non-reachability.
As reachability is recursively enumerable, this provides an algorithm to solve reachability for regular fifo automata.

Let us say that a fifo automaton is \emph{post-regular} (resp. \emph{pred-regular}) if $\post^*(q,w)$ (resp. $\pred^*(q,w)$ ) is regular for all $(q,w) \in Q \times \Sigma^*$. The  fifo automaton is \emph{effectively} post-regular (resp. pred-regular) if the regular language $\post^*(q,w)$ (resp. $\pred^*(q,w)$) is computable. Lossy channel systems are post-regular (since $\post^*(q,w)$ is downward-closed) but they are not effectively post-regular; they are also effectively pred-(upward-closed) regular.

Reachability is (immediately) decidable for both effective post-regular or effective pred-regular fifo automata.
%\alain{montrer que post-regular (pred-regular) sur des (q,w) implique post-regular  (pred-regular) sur des langages rationnels L....ça doit marcher en prenant l'expression rationnelle de L...peut-on montrer que pred-regular n'equivaut pas a post-regular ? contre-exemples ? c'est le cas pour les LCM où pred ok mais pas post}
%
%
%Let us say that a fifo automaton is \emph{regular} if the reachability relation is regular for all $(q,w)$; in this case, the fifo automaton is also both post-regular and pred-regular.
%
Now, we just observe that $\Bad \subseteq \pred^*(\Safe)$ is decidable when $\pred^*(\Safe)$ is a computable regular language and $\Bad$ a regular language:

\begin{theorem}{}
Resilience is decidable for effective pred-regular fifo automata when $\Safe$ and $\Bad$ are regular languages.
\end{theorem}

%		\begin{proof}
%		$\Bad \subseteq \pred^*(\Safe)$ is decidable because $\pred^*(\Safe)$ is a computable regular language.
%		\alain{prover que $\pred^*(\Safe)$ is a computable regular language}
%		\end{proof}
%
%			Since lossy channel systems have (non computable) regular reachability sets, we imediately deduce that.

Since for lossy channel systems $(Q, \Sigma,d,...)$ with $d$ fifo channels and a set of states $L \subseteq Q \times (\Sigma^*)^d$, the set $\pred^*(\uparrow L)$ is a computable upward-closed regular language, we deduce :

\begin{corollary}
Resilience is decidable for lossy channel systems with $\Safe=\uparrow \Safe$ and any regular language $\Bad$.
\end{corollary}

Another proof is possible for LCS since LCS are WSTS with a computable $\uparrow \post^*(S)$ \alain{à prouver et vérifier les autres hypothèses}
%
%%%
\subsection{Resilience for VASS}

>>>>

\section{Resilience for VASS}

### 2023-05-29 definitions.tex
up

	\iffalse
\problemx{Reachability}
{A transition system $\mathscr{S} = (S, \to)$, two configurations $s, t \in S$}
{$s \to^* t$? \\}
	\fi

@@

%We say that $\mathscr{S}$ is {\em upward pre-effective} if it is effective,
%and if there exists an additional Turing machine that computes a finite basis of 
%$\uparrow \pred (\uparrow x)$
%on input $x$, where $x$ is a state of $\mathscr{S}$. 
%
\iffalse
%%%%%%%%%

%%%%%%%%%

\alain{pas ici}
\begin{definition}{\cite{BFM-ic17}} 
A class $C$ of WSTS is {\em ideally effective} if
\begin{itemize}
\item the function mapping the encoding of a configuration $s$ of a WSTS
to the encoding of the ideal $\downarrow s$ is computable;
\item inclusion of ideals is decidable;
\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
\end{itemize}
\end{definition}
%
\fi
%
%
With the ideally effective hypothese, we obtain the decidability of coverability for a class of ordered transition systems larger than WSTS.

@@

%         A downard-closed set $D$ is {\em decidable} if, given $s \in S$, it is decidable whether
%         $s \in D$.

%\textcolor{red}{Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.}
% \textcolor{red}{Do we still demand this ?}
%\mathieu{On peut enlever ce point de discussion ici et en discuter de manière plus complête au moment où on discute de SAFE et BAD plus en détail je pense.}

%
\iffalse
%
\begin{claim}{(stability of upward-closed sets)}
Let $U, V \subseteq S$ be upward-closed. Then the sets $U \cup V$, and $U \cap V$ are upward-closed.
\end{claim}
%
\fi
%

% \begin{claim}
% Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
% \end{claim}

%			\begin{fact}\label{fact basis}
% Fact 3 ([1]). 
% Parosh Aziz Abdulla, Karlis Cerans, Bengt Jonsson & Yih-Kuen Tsay (1996): General Decidability Theorems for Infinite-State Systems. In: Proc. LICS 1996, IEEE Computer Society Press, pp. 313–321,

%For every upward-closed set $U \subseteq S$, there exists a finite basis $B$ of $U$.

% (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
% \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $\min(A)$ est la base minimale}
%			\end{fact}
%

>>>>

@@

With the ideally effective hypothese, we obtain the decidability of coverability for a class of ordered transition systems larger than WSTS.

@@

### 2023-05-29 state-resilience.tex
up

% To convince ourselves that $t$-liveness is undecidable in general for WSTS, we consider the case of Reset VASS. 
The following theorem is a adjustment of Theorem~5.5 from \cite{peterson1981petri} whose proof can be seen in Appendix~\ref{appendix}.

% \alain{tu ne définis pas, même informellement et/ou rapidement, assez les objets que tu utilises: zero reachability problem, Reset VASS}

Adjusting Theorem~5.5 from \cite{peterson1981petri}:

@@

\iffalse
% \mathieu{I moved the proof to the appendix since it's a slight adjustment but takes a lot of room.}
\begin{proof}
See Appendix~\ref{appendix} for the complete proof.
\end{proof}
\fi

Since the zero-reachability problem for Reset $d$-VASS is undecidable, the reduction implies 
% the following:
%
\iffalse
\begin{corollary}
Liveness is undecidable in Reset VASS.
\end{corollary}

Since moreover liveness can be reduced to 
$t$-liveness, we deduce:
\fi
%
%\begin{corollary}
{\sc %$ Reset $d$-VASS 
$t$-liveness} is undecidable.
%\end{corollary}

>>>>

% To convince ourselves that $t$-liveness is undecidable in general for WSTS, we consider the case of Reset VASS. 
The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to Reset VASS, whose proof can be seen in Appendix~\ref{appendix}.

@@

Since the zero-reachability problem for Reset $d$-VASS is undecidable, the reduction implies 
% the following:
%
%
{\sc %$ Reset $d$-VASS 
$t$-liveness} is undecidable.
